---
title: Auto-team Selection, Pre-season
author: Dominic Pearce
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(lpSolve)
library(readr)
library(purrr)
library(tibble)
```

# GW1 Selection

### Using linear programming we set a number of constraints and a variable to optimise our selection on and the computer choses the best possible team based off of last seasons points tallies

Credit to Martin Eastwood at [Pena.lt/y](Pena.lt/y) for the intial script on how to do this!

### Data load

```{r}
df <- read.delim("https://raw.githubusercontent.com/pearcedom/FF/master/16-17/scrapes/16-08-11-scrape.txt") %>% as_tibble
```

### Optimisation variable
Total points are what we really care about here, anything else (goals, assists, etc.) is secondary

```{r}
objective <- df$total_points
```

### Fitting Constraints
Teams must consist of a 15-man squad, 2xGK, 5xDEF, 5xMID and 3xATK. The maximum spend is £100 million. We'll also constrain on chance of playing next round.

*note:* for whatever reason the player data is given in units of 100,000s

```{r}
#position allowances
num_gk <- 2
num_def <- 5
num_mid <- 5
num_fwd <- 3

#binary vectors specifying which rows (i.e. which players) belong to a given position
df$Goalkeeper <- ifelse(df$element_type == "GK", 1, 0)
df$Defender <- ifelse(df$element_type == "DEF", 1, 0)
df$Midfielder <- ifelse(df$element_type == "MID", 1, 0)
df$Forward <- ifelse(df$element_type == "FWD", 1, 0)

likelihood <- df$chance_of_playing_next_round
df$likelihood <- ifelse(is.na(likelihood), 0, ifelse(likelihood >= 75, 0, 1))

#budget
max_cost <- 1000
```

Additionally, a maximum of three players from the same team can be selected

```{r}
team_constraint <- sapply(unique(df$team.y), function(x) ifelse(df$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t
```

### Combine all the information so far into a single matrix
This results in a 505-column (one per player) and 25-row (1 per variable (4 x position, 1 x current cost, 20 x team)) matrix

```{r}
const_mat <- rbind(df$Goalkeeper, df$Defender, df$Midfielder, df$Forward, df$likelihood, df$now_cost, team_constraint)
```

### Lastly for the constraints, we specify their directions and define the righthand-side values

```{r}
const_dir <- c("=", "=", "=", "=","<=", rep("<=", 21))
const_rhs <- c(num_gk, num_def, num_mid, num_fwd, 2, max_cost, rep(3, 20))
#i.e. =2GK, =5DEF, =5MID, =3ATK, =£100m, <=2 injured players, <=three players from a single team
```

### And solve
Here we feed the function
- "max": specifying that we want as much as possible
- "objective": based on total score
- "const_mat": our constraint vectors all housed together
- "const_dir": the direction of our constraints (i.e. whether it's less than/equal to)
- "const_rhs": the righthand-side values (i.e. the value that it must be less than/equal to)
- "all.bin": specifies that all variables are binary
- "all.int": specifies that all variables are integers

```{r}
x <- lp("max", objective, const_mat, const_dir, const_rhs, all.bin=TRUE, all.int=TRUE)
```

### The squad!
```{r}
df_out <- df[which(x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(df_out[order(df_out$element_type),])
```

### We still need to decided on an 11 however, we quickly rework the solution, adding in the following constraints
- 11 men
- no injuries (75% chance to play allowed)
- total cost constraint is removed

```{r}
team_df <- df[which(x$solution==1),]

team_objective <- team_df$total_points

team_team_constraint <- sapply(unique(team_df$team.y), function(x) ifelse(team_df$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

team_const_mat <- rbind(team_df$Goalkeeper, team_df$Defender, team_df$Defender, team_df$Midfielder, team_df$Midfielder, team_df$Forward, team_df$Forward, team_df$likelihood, team_team_constraint, 1)

team_const_dir <- c("=", "<=",  ">=", "<=", ">=", "<=", ">=", "=", rep("<=", length(unique(team_df$team.y))), "=")
team_const_rhs <- c( 1, 5, 3, 5, 2, 3, 1, 0, rep(3, length(unique(team_df$team.y))), 11)

team_x <- lp("max", team_objective, team_const_mat, team_const_dir, team_const_rhs, all.bin=TRUE, all.int=TRUE)

team_df_out <- team_df[which(team_x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(team_df_out[order(team_df_out$element_type),])
```

### Captain Selection
For the first week this will just be based on highest total score last year - **Mahrez** (*Lukaku VC*)
```{r}
team_df_out$second_name[order(team_df_out$total_points, decreasing = TRUE)][1:2]
```

# Post-GW1 Selections

Having completed our original squad each week we can make a substitution, to remove an injured/under performing player. For GW2-4ish we'll continue to use last seasons data, until this seasons information becomes a bit more robust.

To actually achieve this, we can simply add another constraint, that our team must contain >10 of the previous weeks players. This will mean that for the first couple of weeks only injuries are likely to force a line-up change.

```{r}
df_gw2 <- read.delim("https://raw.githubusercontent.com/pearcedom/FF/master/16-17/scrapes/16-08-18-scrape.txt") %>% as_tibble

sub_objective <- df_gw2$total_points

#binary vectors specifying which rows (i.e. which players) belong to a given position
df_gw2$Goalkeeper <- ifelse(df_gw2$element_type == "GK", 1, 0)
df_gw2$Defender <- ifelse(df_gw2$element_type == "DEF", 1, 0)
df_gw2$Midfielder <- ifelse(df_gw2$element_type == "MID", 1, 0)
df_gw2$Forward <- ifelse(df_gw2$element_type == "FWD", 1, 0)

likelihood <- df_gw2$chance_of_playing_next_round
df_gw2$likelihood <- ifelse(is.na(likelihood), 0, ifelse(likelihood >= 75, 0, 1))

sub_team_constraint <- sapply(unique(df_gw2$team.y), function(x) ifelse(df_gw2$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

gw1_picks <- rep(0, nrow(df_gw2))
gw1_picks[map_int(as.character(team_df$web_name), function(x) which(df_gw2$web_name == x))] <- 1

df_gw2$squad <- gw1_picks

sub_const_mat <- rbind(df_gw2$Goalkeeper, df_gw2$Defender, df_gw2$Midfielder, df_gw2$Forward, df_gw2$likelihood, df_gw2$now_cost, sub_team_constraint, df_gw2$squad)
sub_const_dir <- c("=", "=", "=", "=","<=", rep("<=", 21), "=")
sub_const_rhs <- c(num_gk, num_def, num_mid, num_fwd, 2, max_cost, rep(3, 20), 14)

#DT::datatable(rbind(sub_const_dir, sub_const_rhs))

sub_x <- lp("max", sub_objective, sub_const_mat, sub_const_dir, sub_const_rhs, all.bin=TRUE, all.int=TRUE)

sub_out <- df_gw2[which(sub_x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(sub_out[order(sub_out$element_type),])
```

### Check who's been removed

```{r}
sapply(as.character(team_df$second_name), function(x) which(x == as.character(sub_out$second_name)))
```

### And we need to select our starting 11 once again

```{r}
team_df_gw2 <- df_gw2[which(sub_x$solution==1),]

team_objective_gw2 <- team_df_gw2$total_points

team_team_constraint_gw2 <- sapply(unique(team_df_gw2$team.y), function(x) ifelse(team_df_gw2$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

team_const_mat_gw2 <- rbind(team_df_gw2$Goalkeeper, team_df_gw2$Defender, team_df_gw2$Defender, team_df_gw2$Midfielder, team_df_gw2$Midfielder, team_df_gw2$Forward, team_df_gw2$Forward, team_df_gw2$likelihood, team_team_constraint_gw2, 1)

team_const_dir_gw2 <- c("=", "<=",  ">=", "<=", ">=", "<=", ">=", "=", rep("<=", length(unique(team_df_gw2$team.y))), "=")
team_const_rhs_gw2 <- c( 1, 5, 3, 5, 2, 3, 1, 0, rep(3, length(unique(team_df_gw2$team.y))), 11)

team_x_gw2 <- lp("max", team_objective_gw2, team_const_mat_gw2, team_const_dir_gw2, team_const_rhs_gw2, all.bin=TRUE, all.int=TRUE)

team_df_out_gw2 <- team_df_gw2[which(team_x_gw2$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(team_df_out_gw2[order(team_df_out_gw2$element_type),])
```

### Captain Selection
For the first week this will just be based on highest total score last year - **Mahrez** (*Lukaku VC*)
```{r}
team_df_out$second_name[order(team_df_out$total_points, decreasing = TRUE)][1:2]
```

# GW3
```{r}
df_gw3 <- read.delim("https://raw.githubusercontent.com/pearcedom/FF/master/16-17/scrapes/16-08-26-scrape.txt") %>% as_tibble

sub_objective <- df_gw3$total_points

#binary vectors specifying which rows (i.e. which players) belong to a given position
df_gw3$Goalkeeper <- ifelse(df_gw3$element_type == "GK", 1, 0)
df_gw3$Defender <- ifelse(df_gw3$element_type == "DEF", 1, 0)
df_gw3$Midfielder <- ifelse(df_gw3$element_type == "MID", 1, 0)
df_gw3$Forward <- ifelse(df_gw3$element_type == "FWD", 1, 0)

likelihood <- df_gw3$chance_of_playing_next_round
df_gw3$likelihood <- ifelse(is.na(likelihood), 0, ifelse(likelihood >= 75, 0, 1))

sub_team_constraint <- sapply(unique(df_gw3$team.y), function(x) ifelse(df_gw3$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

gw2_picks <- rep(0, nrow(df_gw3))
#whoops, accidentally subbed wrong player last week!
gw2_wrong <- as.character(team_df_gw2$web_name)
gw2_correct <- map_chr(gw2_wrong, function(x) ifelse(x == "Ayew", "Arnautovic", x))
gw2_picks[map_int(gw2_correct, function(x) which(df_gw3$web_name == x))] <- 1

df_gw3$squad <- gw2_picks
df_gw3[df_gw3$squad == 1,c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]


sub_const_mat <- rbind(df_gw3$Goalkeeper, df_gw3$Defender, df_gw3$Midfielder, df_gw3$Forward, df_gw3$likelihood, df_gw3$now_cost, sub_team_constraint, df_gw3$squad)
sub_const_dir <- c("=", "=", "=", "=","<=", rep("<=", 21), "=")
sub_const_rhs <- c(num_gk, num_def, num_mid, num_fwd, 2, max_cost, rep(3, 20), 14)

#DT::datatable(rbind(sub_const_dir, sub_const_rhs))

sub_x <- lp("max", sub_objective, sub_const_mat, sub_const_dir, sub_const_rhs, all.bin=TRUE, all.int=TRUE)

sub_out <- df_gw3[which(sub_x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(sub_out[order(sub_out$element_type),])
```

### Check who's been removed

```{r}
sapply(as.character(team_df$second_name), function(x) which(x == as.character(sub_out$second_name)))
```

### And we need to select our starting 11 once again

```{r}
team_df_gw3 <- df_gw3[which(sub_x$solution==1),]

team_objective_gw3 <- team_df_gw3$total_points

team_team_constraint_gw3 <- sapply(unique(team_df_gw3$team.y), function(x) ifelse(team_df_gw3$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

team_const_mat_gw3 <- rbind(team_df_gw3$Goalkeeper, team_df_gw3$Defender, team_df_gw3$Defender, team_df_gw3$Midfielder, team_df_gw3$Midfielder, team_df_gw3$Forward, team_df_gw3$Forward, team_df_gw3$likelihood, team_team_constraint_gw3, 1)

team_const_dir_gw3 <- c("=", "<=",  ">=", "<=", ">=", "<=", ">=", "=", rep("<=", length(unique(team_df_gw3$team.y))), "=")
team_const_rhs_gw3 <- c( 1, 5, 3, 5, 2, 3, 1, 0, rep(3, length(unique(team_df_gw3$team.y))), 11)

team_x_gw3 <- lp("max", team_objective_gw3, team_const_mat_gw3, team_const_dir_gw3, team_const_rhs_gw3, all.bin=TRUE, all.int=TRUE)

team_df_out_gw3 <- team_df_gw3[which(team_x_gw3$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(team_df_out_gw3[order(team_df_out_gw3$element_type),])
```


### **MUST ACTUALLY FIX THIS PROPERLY**
# GW4
```{r}
df_gw4 <- read.delim("~/Documents/FF/16-17/scrapes/16-09-10.txt") %>% as_tibble

sub_objective <- df_gw4$total_points

#binary vectors specifying which rows (i.e. which players) belong to a given position
df_gw4$Goalkeeper <- ifelse(df_gw4$element_type == "GK", 1, 0)
df_gw4$Defender <- ifelse(df_gw4$element_type == "DEF", 1, 0)
df_gw4$Midfielder <- ifelse(df_gw4$element_type == "MID", 1, 0)
df_gw4$Forward <- ifelse(df_gw4$element_type == "FWD", 1, 0)

likelihood <- df_gw4$chance_of_playing_next_round
df_gw4$likelihood <- ifelse(is.na(likelihood), 0, ifelse(likelihood >= 75, 0, 1))

sub_team_constraint <- sapply(unique(df_gw4$team.y), function(x) ifelse(df_gw4$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

gw3_picks <- rep(0, nrow(df_gw4))
gw3_picks[map_int(as.character(sub_out$second_name), function(x) which(df_gw4$web_name == x))] <- 1

df_gw4$squad <- gw3_picks
df_gw4[df_gw4$squad == 1,c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]


sub_const_mat <- rbind(df_gw4$Goalkeeper, df_gw4$Defender, df_gw4$Midfielder, df_gw4$Forward, df_gw4$likelihood, df_gw4$now_cost, sub_team_constraint, df_gw4$squad)
sub_const_dir <- c("=", "=", "=", "=","<=", rep("<=", 21), "=")
sub_const_rhs <- c(num_gk, num_def, num_mid, num_fwd, 2, max_cost, rep(3, 20), 14)

#DT::datatable(rbind(sub_const_dir, sub_const_rhs))

sub_x <- lp("max", sub_objective, sub_const_mat, sub_const_dir, sub_const_rhs, all.bin=TRUE, all.int=TRUE)

sub_out <- df_gw4[which(sub_x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(sub_out[order(sub_out$element_type),])
```

### Check who's been removed

```{r}
sapply(as.character(team_df$second_name), function(x) which(x == as.character(sub_out$second_name)))
```

### And we need to select our starting 11 once again

```{r}
team_df_gw4 <- df_gw4[which(sub_x$solution==1),]

team_objective_gw4 <- team_df_gw4$total_points

team_team_constraint_gw4 <- sapply(unique(team_df_gw4$team.y), function(x) ifelse(team_df_gw4$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

team_const_mat_gw4 <- rbind(team_df_gw4$Goalkeeper, team_df_gw4$Defender, team_df_gw4$Defender, team_df_gw4$Midfielder, team_df_gw4$Midfielder, team_df_gw4$Forward, team_df_gw4$Forward, team_df_gw4$likelihood, team_team_constraint_gw4, 1)

team_const_dir_gw4 <- c("=", "<=",  ">=", "<=", ">=", "<=", ">=", "=", rep("<=", length(unique(team_df_gw4$team.y))), "=")
team_const_rhs_gw4 <- c( 1, 5, 3, 5, 2, 3, 1, 0, rep(3, length(unique(team_df_gw4$team.y))), 11)

team_x_gw4 <- lp("max", team_objective_gw4, team_const_mat_gw4, team_const_dir_gw4, team_const_rhs_gw4, all.bin=TRUE, all.int=TRUE)

team_df_out_gw4 <- team_df_gw4[which(team_x_gw4$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(team_df_out_gw4[order(team_df_out_gw4$element_type),])
```

# GW5
## Wildcard time!

### Data load

```{r}
df <- read_tsv("https://raw.githubusercontent.com/pearcedom/FF/master/16-17/scrapes/16-09-16.txt")
```

### Optimisation variable
Total points are what we really care about here, anything else (goals, assists, etc.) is secondary

```{r}
objective <- df$total_points
```

### Fitting Constraints
Teams must consist of a 15-man squad, 2xGK, 5xDEF, 5xMID and 3xATK. The maximum spend is £100 million. We'll also constrain on chance of playing next round.

*note:* for whatever reason the player data is given in units of 100,000s

```{r}
#position allowances
num_gk <- 2
num_def <- 5
num_mid <- 5
num_fwd <- 3

#binary vectors specifying which rows (i.e. which players) belong to a given position
df$Goalkeeper <- ifelse(df$element_type == "GK", 1, 0)
df$Defender <- ifelse(df$element_type == "DEF", 1, 0)
df$Midfielder <- ifelse(df$element_type == "MID", 1, 0)
df$Forward <- ifelse(df$element_type == "FWD", 1, 0)

likelihood <- df$chance_of_playing_next_round
df$likelihood <- ifelse(is.na(likelihood), 0, ifelse(likelihood >= 75, 0, 1))

#budget
max_cost <- 991 #oops, money has been lost by transfers!
```

Additionally, a maximum of three players from the same team can be selected

```{r}
team_constraint <- sapply(unique(df$team.y), function(x) ifelse(df$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t
```

### Combine all the information so far into a single matrix
This results in a 505-column (one per player) and 25-row (1 per variable (4 x position, 1 x current cost, 20 x team)) matrix

```{r}
const_mat <- rbind(df$Goalkeeper, df$Defender, df$Midfielder, df$Forward, df$likelihood, df$now_cost, team_constraint)
```

### Lastly for the constraints, we specify their directions and define the righthand-side values

```{r}
const_dir <- c("=", "=", "=", "=","<=", rep("<=", 21))
const_rhs <- c(num_gk, num_def, num_mid, num_fwd, 2, max_cost, rep(3, 20))
#i.e. =2GK, =5DEF, =5MID, =3ATK, =£100m, <=2 injured players, <=three players from a single team
```

### And solve
Here we feed the function
- "max": specifying that we want as much as possible
- "objective": based on total score
- "const_mat": our constraint vectors all housed together
- "const_dir": the direction of our constraints (i.e. whether it's less than/equal to)
- "const_rhs": the righthand-side values (i.e. the value that it must be less than/equal to)
- "all.bin": specifies that all variables are binary
- "all.int": specifies that all variables are integers

```{r}
x <- lp("max", objective, const_mat, const_dir, const_rhs, all.bin=TRUE, all.int=TRUE)
```

### The squad!
```{r}
df_out <- df[which(x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(df_out[order(df_out$element_type),])
```

### We still need to decided on an 11 however, we quickly rework the solution, adding in the following constraints
- 11 men
- no injuries (75% chance to play allowed)
- total cost constraint is removed

```{r}
team_df <- df[which(x$solution==1),]

team_objective <- team_df$total_points

team_team_constraint <- sapply(unique(team_df$team.y), function(x) ifelse(team_df$team.y == x, 1, 0), USE.NAMES = TRUE) %>% as_tibble %>% t

team_const_mat <- rbind(team_df$Goalkeeper, team_df$Defender, team_df$Defender, team_df$Midfielder, team_df$Midfielder, team_df$Forward, team_df$Forward, team_df$likelihood, team_team_constraint, 1)

team_const_dir <- c("=", "<=",  ">=", "<=", ">=", "<=", ">=", "=", rep("<=", length(unique(team_df$team.y))), "=")
team_const_rhs <- c( 1, 5, 3, 5, 2, 3, 1, 0, rep(3, length(unique(team_df$team.y))), 11)

team_x <- lp("max", team_objective, team_const_mat, team_const_dir, team_const_rhs, all.bin=TRUE, all.int=TRUE)

team_df_out <- team_df[which(team_x$solution==1), c("first_name", "second_name", "now_cost", "team.y", "element_type", "total_points")]

knitr::kable(team_df_out[order(team_df_out$element_type),])
```

### Captain Selection
For the first week this will just be based on highest total score last year - **Mahrez** (*Lukaku VC*)
```{r}
team_df_out$second_name[order(team_df_out$total_points, decreasing = TRUE)][1:2]
````

